---
title: "Quantifying effort and distribution of PSPAP gears"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:  
  html_document:
    theme: readable
    highlight: tango
---

<!--
rmarkdown::render("index.Rmd")# build page
-->

```{r,echo=FALSE,warning=FALSE, message=FALSE}
source("R/1_global.R")
source("R/2_functions.R")
source("R/3_load-and-clean.R")
source("R/4_figures.R")
source("R/5_tables.R")
source("R/6_analysis.R")
```  
## Objectives

1. Characterize temporal use of gear for PSPAP duration
2. Quantify effort and associated dispersion for each PSPAP gear

## Methods

* fit distribution to effort for common gear types

## Results





## Figures

```{r,echo=FALSE}
figures(1)
```

Figure 1. Illustration of gears used since 2003 for Pallid Sturgeon 
Population assessment in the Missouri River. Top panel is for Lower basin (RPMA 4)
and bottom panel is for Upper basin (RPMA 2). 

## Tables

Table 1. Summary of frequency of use for each gear type by year. 
```{r,echo=FALSE}
library(knitr)
kable(tables(1),
    col.names=c("Basin","PSPAP gear code", "Gear id",2003:2016))
    
```