---
title: Effect of violating closure in a closed population
output: 
  html_document:
    includes:
      in_header: header_banner.html 
bibliography: PSref.bib
csl: ecology.csl  
---

<!---
rmarkdown::render_site("2017-06-XX-RD-and-closure.Rmd")# build website
--->


### In a nutshell

### Engage in the process

### Background




```{r}
library(rowr)
## make a ZZ nrow=inds, ncol=nprim*nsec
## PRIMARY OCCASIONS
nprim = 10

## SURVIVAL
phi=rep(0.8,nprim)
# FECUNDITY
f=rep(0.1,nprim)
n=1000 # initial population size
n_inn=300 # initial number in study area


## RANDOM MOVEMENT PRIMARY
gam_prime=rep(0.7,nprim) # UNOBSERVABLE[t-1]--> UNOBSERVABLE[t]
gam_d_prime=1-gam_prime # OBSERVABLE[t-1] --> UNOBSERVABLE[t]


## SECONDARY OCCASIONS
nsec=rep(4,nprim)
## RANDOM MOVEMENT SECONDARY

gam_d_prime2<-rep(0.1,nprim) # OBSERVABLE[t-1] --> UNOBSERVABLE[t]
gam_prime2<- 1-gam_d_prime2# UNOBSERVABLE[t-1]--> UNOBSERVABLE[t]

## CAPTURE PROBABILITY
p=rep(0.3,nprim)
    
    
## SUPER POPULATION
Z<- matrix(0,n,nprim)
Z_inn<-Z
Z_out<-Z

Z[,1]<-1
Z_inn[1:n_inn,1]<-1
Z_out[,1]<- 1*(1-Z_inn[,1])

# SIMULATE CAPTURE HISTORY
## MORE THAN 1 CAPTURE OCCASIONS
ch<- matrix(0,nrow(Z),sum(nsec))

## LINK SECONDARY OCCASION TO PRIMARY OCCASION
primsec<- matrix(
    cbind(rep(1:nprim,nsec),
    c(1:sum(nsec))), ncol=2)

 
## CAPTURE HISTORY FOR T=1    
## UPDATE WHO IS IN AND OUT
## AFTER DAY TO DAY MOVEMENT
i<-1       
ch<-Z_inn2<-Z_out2<- c()
for(k in 1:nsec[i])
    {
    ch<- cbind(ch,rbinom(nrow(Z_inn),1,p[i]*Z_inn[,i]*Z[,i]))    
    ## MOVEMENT OUT OF STUDY AREA
    innToOut<- rbinom(n=nrow(Z),
        size=1,
        prob=gam_d_prime2[i]*Z[,i]*Z_inn[,i])
    innToInn<- (1-innToOut)*(Z[,i]*Z_inn[,i])
    ## REMAINING OUT OF STUDY AREA
    outToOut<- rbinom(n=nrow(Z),
        size=1,
        prob=gam_prime2[i]*Z[,i]*Z_out[,i])
    outToInn<-  (1-outToOut)*(Z[,i]*Z_out[,i])
    Z_inn[,i]<-0 
    Z_inn[which(innToInn==1 | outToInn==1),i]<- 1
    Z_out[,i]<-0 
    Z_out[which(outToOut==1 | innToOut==1),i]<- 1
    Z_inn2<- cbind(Z_inn2,Z_inn[,i])
    Z_out2<- cbind(Z_out2,Z_out[,i])
    }    

 

## SIMULATE POPULATION DYNAMICS
for(i in 2:nprim)
    {
    ## SURVIVAL
    Z[,i]<- rbinom(n=nrow(Z),
        size=1,
        prob=phi[i-1]*Z[,i-1])
    
    ## MOVEMENT OUT OF STUDY AREA
    innToOut<- rbinom(n=nrow(Z),
        size=1,
        prob=gam_d_prime[i-1]*Z[,i]*Z_inn[,i-1])
    innToInn<- (1-innToOut)*(Z[,i]*Z_inn[,i-1])
    
    ## REMAINING OUT OF STUDY AREA
    outToOut<- rbinom(n=nrow(Z),
        size=1,
        prob=gam_prime[i-1]*Z[,i]*Z_out[,i-1])
    outToInn<-  (1-outToOut)*(Z[,i]*Z_out[,i-1])
    Z_inn[,i]<-0 
    Z_inn[which(innToInn==1 | outToInn==1),i]<- 1
    Z_out[,i]<-0 
    Z_out[which(outToOut==1 | innToOut==1),i]<- 1

   
    ## SIMULATE DAY TO DAY MOVEMENT AND
    ## CAPTURE HISTORY    
    for(k in 1:nsec[i])
        {
        ch<- cbind.fill(ch,rbinom(nrow(Z_inn),1,p[i]*Z_inn[,i]*Z[,i]))        
        ## MOVEMENT OUT OF STUDY AREA
        innToOut<- rbinom(n=nrow(Z),
            size=1,
            prob=gam_d_prime2[i]*Z[,i]*Z_inn[,i])
        innToInn<- (1-innToOut)*(Z[,i]*Z_inn[,i])
        ## REMAINING OUT OF STUDY AREA
        outToOut<- rbinom(n=nrow(Z),
            size=1,
            prob=gam_prime2[i]*Z[,i]*Z_out[,i])
        outToInn<-  (1-outToOut)*(Z[,i]*Z_out[,i])
        Z_inn[,i]<-0 
        Z_inn[which(innToInn==1 | outToInn==1),i]<- 1
        Z_out[,i]<-0 
        Z_out[which(outToOut==1 | innToOut==1),i]<- 1
        Z_inn2<- cbind.fill(Z_inn2,Z_inn[,i])
        Z_out2<- cbind.fill(Z_out2,Z_out[,i])
        }  
    
    ## RECRUITMENT
    ## ASSUME RECRUITS SETTLE OUTSIDE OF 
    ## STUDY AREA
    R<-rpois(1,sum(Z[,i])*f[i-1])
    app<- matrix(0,R,nprim)
    app[,i]<-1
    Z<- rbind(Z,app)
    Z_out<- rbind(Z_out,app)
    app[]<-0
    Z_inn<- rbind(Z_inn,app)
    }


    
    

    
## SUBSET OUT FISH THAT ARE NEVER CAPTURED
ch<- ch[which(apply(ch,1,sum)!=0),]
## NEED THIS FOR RMARK
# prep data for processing
ch<- data.frame(ch=apply(ch,1,paste0,collapse=""),
    freq=1,stringsAsFactors=FALSE)
    
## SET UP TIME INTERRVALS
## FOR PROGRAM MARK
ends<-cumsum(nsec) # last sampling occasion
occs<- rep(0,sum(nsec))
occs[ends]<-1# last occasion in primary
occs<- occs[-length(occs)]# drop last 1 for processing 


out<-(list(ch=ch,
    occs=occs,
    nprim=nprim,
    nsec=nsec,
    Z=Z,
    Z_inn=Z_inn,
    Z_out=Z_out))

    

    
    library(RMark)
rd<-process.data(data=out$ch, 
    model="Robust", 
    time.intervals=out$occs)
S=list(formula=~1)# SURVIVAL
# SHARE = TRUE TO SET C = P
p=list(formula=~1,share=TRUE)# CAPTURE PROBABILITY
f0<- list(formula=~time) # NUMBER NOT ENCOUNTERED
GammaDoublePrime=list(formula=~1,share=TRUE)
GammaPrime=list(formula=~1)
fit<-mark(data = rd, 
	model = "Robust", 
    time.intervals=time.intervals,
	model.parameters=list(
		S=S,
		GammaDoublePrime=GammaDoublePrime,
		# GammaPrime=GammaPrime, # not needed when share=TRUE
		p=p),
	threads=2,
	brief=TRUE)
derived<- fit$results$derived$`N Population Size`
matplot((cbind(colSums(out$Z_inn),derived$estimate)),type='b')

#RDPdfClosed

plogis(summary(fit)$beta$estimate)[1:3]
#f0 = beta[1] + effect of session then exp
#exp(summary(fit)$beta$estimate)[-c(1:3)]
#unlist(summary(fit)$real$f0)



## Notes
* Recruitment does not effect estimates of phi, gamma,
and p












## Introduction

Relevant assumptions 

1. Fish do not migrate into or out of the study area
2. Fish vulnerable to capture over the study period

### Objectives

1. Evaluate use of Robust Design in broodstock abundance estimates
2. Quantify effects of violating the assumption that the population
is closed to emigration.

## Methods

### Simulating a virtual population

#### Initial abundance

```{r}
n<- 1000
n_unobservable=2000
nprim=5
nsec=3
phi=0.8
gamma_prime=0.3
gamma_dblprime=0.3
p=0.3

    
super_n<- n+n_unobservable # SUPER POPULATION (ONSITE + OFFSITE)
Z<- matrix(0,super_n,nprim)
Z_observable<- matrix(0,super_n,nprim)
Z_unobservable<- matrix(0,super_n,nprim)
Z[1:super_n,1]<-1
Z_observable[1:n,1]<-1
Z_unobservable[(n+1):(n+n_unobservable),1]<-1


# SURVIVAL, EMIGRATION, AND MIGRATION
for(i in 1:super_n) # FOR EACH INDIVIDUAL IN THE SUPER POPULATION
    {
    for(j in 2:(nprim)) # FOR PRIMARY OCCASIONS 2:NPRIM
        {
        # DOES A FISH SURVIVE, IF IT WAS PREVIOUSLY ALIVE?
        Z[i,j]<- rbinom(1,1,phi[j-1])*Z[i,j-1]

        prs<-c(
            # OBSERVABLE[t-1] --> UNOBSERVABLE[t]
            (1-Z_unobservable[i,j-1])*(gamma_dblprime[j-1]),
            
            # OBSERVABLE[t-1] --> OBSERVABLE[t]
            (1-Z_unobservable[i,j-1])*(1-gamma_dblprime[j-1]),
            
            # UNOBSERVABLE[t-1]--> UNOBSERVABLE[t]
            Z_unobservable[i,j-1]*gamma_prime[j-1],
            
            # UNOBSERVABLE[t-1]--> 	OBSERVABLE[t]
            Z_unobservable[i,j-1]*(1-gamma_prime[j-1]))

        state<-sample(1:4,1,prs,replace=FALSE)
        if(state==1 & Z[i,j]==1){Z_unobservable[i,j]<- 1}
        if(state==2 & Z[i,j]==1){Z_observable[i,j]<-1}
        if(state==3 & Z[i,j]==1){Z_unobservable[i,j]<-1}
        if(state==4 & Z[i,j]==1){Z_observable[i,j]<-1}
        }
    }
    
    
    
        
    # SIMULATE CAPTURE HISTORY
    ## MORE THAN 1 CAPTURE OCCASIONS
    ch<- matrix(0,nrow(Z),sum(nsec))
    avail2Capture<-ch
    ## LINK SECONDARY OCCASION TO PRIMARY OCCASION
    primsec<- matrix(cbind(rep(1:nprim,nsec),
        c(1:sum(nsec))), ncol=2)
    ## CAPTURE HISTORY CONDITIONAL ON BEING PRESENT (I.E., NOT EMIGRATED)
    for(i in 1:super_n)
        {
        for(j in 1:sum(nsec))
            {
            primocc<- primsec[j,1]
            avail2Capture[i,j]<-Z_observable[i,primsec[j,1]]
            ch[i,j]<- rbinom(1,1,p[primocc]*avail2Capture[i,j])
            }
        }
        
    ## SUBSET OUT FISH THAT ARE NEVER CAPTURED
    ch<- ch[which(apply(ch,1,sum)!=0),]

    ## NEED THIS FOR RMARK
    ch2<- data.frame(ch=apply(ch,1,paste0,collapse=""),freq=1,stringsAsFactors=FALSE)# prep data for processing

    # NEED TO MAKE A MATRIX OF THAT INDICATES THE FIRST OCCASION
    ends<-cumsum(nsec) # last sampling occasion
    occs<- rep(0,sum(nsec))
    occs[ends]<-1# last occasion in primary
    occs<- occs[-length(occs)]# drop last 1 for processing 


    return(list(ch=ch2,occs=occs))        
    }

    
    
nprim<- 10
nsec<- floor(runif(nprim,5,10))
phi	<- rep(0.8,nprim-1)

# CAPTURE PROBABILITY
p	<- rep(0.3,nprim)

gamma_prime	<- rep(0.3, nprim-1) # pr(unobservable @ t-1  --> unobservable @ t )
gamma_dblprime<- rep(0.4, nprim-1) # pr(observable @ t-1 --> unobservable  @ t   )

n<- 175 # initial population size
n_unobservable<- 200 # UNOBSERVABLE

out<-sim_rd_ch(n=n, 
    n_unobservable=n_unobservable, 
    nprim=nprim,
    nsec=nsec,
    phi=phi,
    gamma_prime=gamma_prime,
    gamma_dblprime=gamma_dblprime,
    p=p) 


rd<-process.data(data=out$ch, model="Robust", time.intervals=out$occs)
rd_ddl<-make.design.data(rd)

S=list(formula=~1)# SURVIVAL
# SHARE = TRUE TO SET C = P
p=list(formula=~1,share=TRUE)# CAPTURE PROBABILITY
f0<- list(formula=~1) # NUMBER NOT ENCOUNTERED
GammaDoublePrime=list(formula=~1)
GammaPrime=list(formula=~1)
fit<-mark(data = rd, 
	model = "Robust", 
    time.intervals=time.intervals,
	model.parameters=list(
		S=S,
		GammaDoublePrime=GammaDoublePrime,
		GammaPrime=GammaPrime,
		p=p),
	threads=2,
	brief=TRUE)
derived<- fit$results$derived$`N Population Size`
saveRDS(derived,"./output/derived.RDS")

cleanup(ask=FALSE)





#### Recruitment



#### Survival

#### Migration rates

Assumes random movement where $\gamma^{\prime}=\gamma^{\prime\prime}.

see http://www.montana.edu/rotella/documents/502/lab09Robust.Rmd
to specify gamma parameters

Process the data for estimation

```{r}
time.intervals=c(0,0,1,0,0,1,0,0,1,0,0,1,0,0)

rd=process.data(data=rd.inp, model="Robust", time.intervals=time.intervals)

names(rd)

# take a look at the data
head(rd$data)

# check the number of primary occasions 
rd$nocc

# check the number of secondary occasions 
rd$nocc.secondary
```


Set up Gamma'' to vary among primary occasions?
Then specify the Gammas to be shared (random movement)
```{r}
GammaDoublePrime.random=list(formula=~time,share=TRUE)
```

Now we can set up the design matrix to be estimated by Program Mark.

```{r}
fit_random=mark(data=rd, design.parameters=
    list(GammaDoublePrime=list(time.bins=c(1,2,3,5)),
         GammaPrime=list(time.bins=c(2,3,5))),
    right=FALSE, 
    model.parameters=list(S=S.time,
        p=p.session,
        GammaDoublePrime=GammaDoublePrime.random))
```

### Catch process


### Estimating abundance and demographic rates

### Consequence of removing fish on estimates

### Consequences of violating closure assumption



## Results


## Discussion


