



```{r}
source("_R/1_global.R")
source("_R/2_functions.R")
source("_R/3_load-and-clean.R")
source("_R/4_figures.R")
source("_R/5_tables.R")
```




```{r}
# GENERATE THE REFERENCE POPULATION
input<-list()
input$segs<- c(1,2,3,4,7,8,9,10,13,14)
# BENDS DATAFRAME
input$bends=bends
# FISH DENSITY PER RKM
input$fish_density=init_dens
input$nyears<- 10
input$beta0<- 2.9444
input$phi<-matrix(plogis(input$beta0),
	length(input$segs),
	input$nyears-1)# MATRIX OF YEAR TO YEAR AND SEGEMENT SPECIFIC SURVIVALS
input$Linf=NULL
input$k = NULL
input$vbgf_vcv=NULL # VARIANCE COVARIANCE MATRIX
input$initial_length=NULL
input$mv_beta0=NULL
input$mv_beta1=NULL
```

The chunk below simulates 200 populations given the inputs above. 


```{r,eval=FALSE}
ptm <- proc.time()
dir<- ifelse(pcname=="WF-FGNL842","E:/_output","_output")
for(reps in 1:200)
	{
	sim_pop<-reference_population(inputs=input)
	# CRUNCH IS WF-FGNL842
	fn<- paste0(dir,"/1-populations/sim-pop-",reps,".rds")
	saveRDS(sim_pop,file=fn)	
	}
proc.time() - ptm
```



Code below loops over each simulated population and 
saves the catch data.  Took ~230 minutes to do 200 
populations  
  
  
```{r}
pop_dir<- ifelse(pcname=="WF-FGNL842","E:/_output/1-populations/","_output/1-populations/")
pop_dir_out<- ifelse(pcname=="WF-FGNL842","E:/_output/2-catch/","_output/2-catch/")
pops<-dir(pop_dir)

## MAKE CATCH DATA REPLICATES FOR RANDOM DRAWS OF CATCHABILITY AND B0_SD
### PICK SAMLING TYPE
#samp_type="r"
samp_type="f"

### RUN
ptm<-proc.time()
for(reps in 1:length(pops))
    {
    fp<- paste0(pop_dir,pops[reps])
    sim_pop<- readRDS(fp)
    input$id<-sim_pop$inputs$id
    ## MEAN CATCHABILITY
    q_mean<-c(runif(5,0.000000, 0.00005), runif(1,0.00005, 0.001),
              runif(2,0.000000, 0.00005),runif(1,0.00005, 0.001))
    # Accounts for differences in "OT16" and "TN" efforts.
    ## B0_SD  
    B0_sd<-runif(9,0,1.5)
    
    input<-list()
    input$samp_type=samp_type
    input$gears=c("GN14", "GN18", "GN41", "GN81",
            "MF", "OT16", "TLC1", "TLC2", "TN")
    input$catchability=q_mean
    input$B0_sd=B0_sd
    input$effort=effort
    input$deployments=rep(8,9)
    input$occasions=3

    
    ## SAMPLING & CATCH DATA
    dat<-catch_data(sim_pop=sim_pop,
        inputs=input)
    saveRDS(dat,
        file=paste0(pop_dir_out,"catch-",reps,".rds"))  
    }
proc.time()-ptm
```


        
 





 
        
# SIMULATE EFFORT & CATCH DATA FOR A FIXED B0_SD GRID
## MEAN CATCHABILITY

proc.time()-ptm

dat<-samp_dat(sim_pop=sim_pop,
    catchability=rep(0.00001,9),
    B0_sd=runif(9, 0,1.5),
    effort=effort,
    occasions=3)
saveRDS(dat,file="output/test.Rds") 



proc.time()-ptm
#user       system    elapsed 
#15795.57   68.90     15988.57
```

