---
title: "Tables"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "Missouri River Pallid Sturgeon Technical Team"
output:  
  html_document:
    theme: readable
    highlight: tango
  pdf_document:
    highlight: zenburn 
  word_document:
    highlight: monochrome
    fig_width: 6.5  
    fig_height: 6.5 
    reference_docx: tmp-editing.docx
---

<!--
rmarkdown::render("_Report-5-03-appendix",output_format='word_document')# build page
-->



Box 1. Function used to simulate reference populations given pallid 
sturgeon density, survival, and spatial structure. 


```{r,echo=FALSE}
print(reference_population)
```


##### page break



Box 1. Function used to simulate which bends are sampled each year given 
a sampling strategy. 

```{r,echo=FALSE}
print(bend_samples)
```

##### page break

Box 2. Function used to simulate the catch data given a sampling 
strategy. Note, this function calls the function "bend_samples", which 
is the function defined in Box 1. 


```{r,echo=FALSE}
print(catch_data)
```
