---
title: "Tables"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "Missouri River Pallid Sturgeon Technical Team"
output:  
  html_document:
    theme: readable
    highlight: tango
  pdf_document:
    highlight: zenburn 
  word_document:
    highlight: monochrome
    fig_width: 6.5  
    fig_height: 6.5 
    reference_docx: tmp-editing.docx
---

<!--
rmarkdown::render("_Report-5-02-tables",output_format='word_document')# build page
-->

```{r, warning=FALSE, include=FALSE}
source("_R/1_global.R")
source("_R/2_functions.R")
#effort_data<-TRUE   # NEEDED TO CREATE TABLE 4 FROM SCRATCH...COMMENT THIS OUT TO SAVE TIME
source("_R/3_load-and-clean.R")
source("_R/4_figures.R")
source("_R/5_tables.R")
```



Table 1. Summary of bends within PSPAP Missouri river segments.


```{r,echo=FALSE,eval=TRUE}
tbl<-tables(1)
tbl$basin<- ifelse(tbl$basin=="lower","LB","UB")
tbl$b_segment<-as.numeric(tbl$b_segment)
tbl<- tbl[order(tbl$b_segment),]
tbl[duplicated(tbl$basin),]$basin<-""
kable(tbl,row.names=FALSE,
    col.names=c("Basin","Segment","Number of bends",
        "Minimum length (km)", "Mean length (km)","Maximum length (km)"),
    align="lccccc")
```



##### Page Break

Table 2. Segment and river level densities by origin: hatchery (H) and 
natural/wild (W).  Minimum and maximum densities were taken from data 
across a few recent years (when available), while mean densities are 
those reported in the literature from the most recent year's data.   

```{r,warning=FALSE, echo=FALSE,eval=TRUE}
tbl2<-tables(2)
tbl2$max_year[which(tbl2$max_year==2008)]<-"2008*"
tbl2$rpma<-ifelse(tbl2$rpma==2, "UB", "LB")
tbl2[duplicated(tbl2$rpma),]$rpma<-""
tbl2[duplicated(tbl2$segments),]$segments<-""
kable(tbl2,row.names=FALSE,
    col.names=c("Basin","Segments", "Fish Type", "Minimum Density (fish/rkm)",
                "Maximum Density (fish/rkm)", "Mean Density (fish/rkm)", 
                "Most Recent Year", "References"),
    align="llcccccc")
```
  
*Estimated year of data collection based on reference date.    



##### page break


Table 3. Number of Bends Sampled per Segment
```{r, echo=FALSE,eval=TRUE}
no_bends<-data.frame(Segment=c(2:4,7:10,13,14), 
    No_Bends=c(12, 21, 12, 12, 15, 20, 10, 11, 14))
kable(no_bends, row.names=FALSE, 
      col.names=c("Segment", "No. of Sampled Bends"))
```

##### page break

Table 4. Summary of effort data by gear and basin, where effort is 
measured in minutes of gear use per deployment. The shape and rate columns are the results of 
fitting a gamma distribution to the data. 


```{r,warning=FALSE,echo=FALSE,eval=TRUE}
#tbl<-tables_effort(n=4, dat=effort)
#write.csv(tbl,"_output/effortAndGearSummary.csv",row.names=FALSE)
tbl<-read.csv("_output/effortAndGearSummary.csv") 
kable(tbl,row.names=FALSE,
    col.names=c("Basin","PSPAP Gear Code","Gear ID", "No. of Observations", 
        "Mean Effort",
        "SD of Effort", "Minimum Effort", "Maximum Effort", 
        "Median Effort","Shape", "Rate"),
    align="lccccc")
```



Table X.1. Field office affiliations and Missouri River segments
assigned for PSPAP sampling.

| Field office                             	| Segment          	|
|------------------------------------------	|------------------	|
| Montana Fish Wildlife and Parks (MT)     	| 1, 2, 3          	|
| Missouri River FWCO (MR)                 	| 4                	|
| Great Plains FWCO (GP)                   	| 5, 6             	|
| South Dakota Game Fish and Parks (SD)    	| 7                	|
| Nebraska Game and Parks Commission (NE)  	| 8, 1/2 of 9      	|
| Missouri Department of Conservation (MO) 	| 1/2 of 9, 10, 11 	|
| Columbia FWCO (CF)                       	| 13, 14           	|



Table X.2.  <!--SARA CAN'T COMPILE THIS TABLE...WHAT IS tmp?-->

```{r,echo=FALSE,as.is=TRUE,eval=TRUE}
source("_R/6-analysis-cost.R")
knitr::kable(tmp,
    row.names=FALSE,
    col.names=c("Field crew", "Mean cost","Median cost", "Std. deviation",
        "Minimum cost","Maximum cost"),align=c("lccccc"),
    digits=2)
```

Table x.3.

Missouri River segments and RPMAs.

| RPMA | Segment | Upstream Rkm | Downstream Rkm | Length (Km) | Description                                                          |
|:----:|:-------:|:------------:|:--------------:|:-----------:|----------------------------------------------------------------------|
|   2  |    1    |     2851     |     2832.4     |     18.6    | Fort Peck Dam downstream to the Milk River Confluence                |
|   2  |    2    |    2832.4    |     2737.5     |     94.9    | Milk River Confluence downstream to Wolf Point, Montana              |
|   2  |    3    |    2737.5    |      2546      |    191.5    | Wolf Point, Montana downstream to the Yellowstone River Confluence   |
|   2  |    4    |     2546     |     2523.45    |    22.55    | Yellowstone River Confluence to the headwaters of Lake Sakakawea     |
|   3  |    5    |    1416.2    |     1359.9     |     56.3    | Fort Randall Dam to the Niobrara River Confluence                    |
|   3  |    6    |    1359.9    |      1327      |     32.9    | Niobrara River/ Confluence to the headwaters of Lewis and Clark Lake |
|   4  |    7    |    1305.2    |      1207      |     98.2    | Gavins Point Dam to Ponca, Nebraska                                  |
|   4  |    8    |     1207     |      958.4     |    248.6    | Ponca, Nebraska to the Platte River Confluence                       |
|   4  |    9    |     958.4    |      591.4     |     367     | Platte River Confluence to the Kansas River Confluence               |
|   4  |    10   |     591.4    |      402.3     |    189.1    | Kansas River Confluence downstream to the Grand River Confluence     |
|   4  |    13   |     402.3    |      209.2     |    193.1    | Grand River to Osage River, Missouri (confluence)                    |
|   4  |    14   |     209.2    |        0       |    209.2    | Osage River to the mouth, Missouri     


## Effort Tables

Table 1. Summary of frequency of use for each gear type by year. An 
asterisk (*) following a gear name indicates that the data from this 
gear was used in the effort analysis (summarized in Table 4). 

```{r,echo=FALSE,eval=FALSE}
## NEED TO FIX BASIN IN THIS ONE.
kable(tables_effort(1,dat=effort),
    col.names=c("Basin","PSPAP gear code", "Gear id",2003:2016),
    row.names=FALSE)
   
```


Table 2. Frequency of use (number of deployments) for each standard gear 
type by bend (within river segment) and year. An asterisk (*) following 
a gear name indicates that the data from this gear was used in the 
effort analysis (summarized in Table 4). 

```{r,echo=FALSE,eval=FALSE}
## NEED TO FIX STANDARD GEAR HERE
kable(tables_effort(2,dat=effort),
    col.names=c("Basin","Segment","Bend","PSPAP gear code", "Gear id",2003:2016),
    row.names =FALSE)
   
```

Table 3. Summary of the number of deployments for each standard gear 
type by bend (within river segment) per year. An asterisk (*) following 
a gear name indicates that the data from this gear was used in the 
effort analysis (summarized in Table 4). 

```{r,echo=FALSE,eval=FALSE}
## NEED TO FIX STANDARD GEAR HERE
kable(tables_effort(3,dat=effort),
    col.names=c("Basin","Segment","Bend","PSPAP gear code", "Gear id", "Minimum","Maximum","Mean","Median", "Total Deployments"),
    row.names =FALSE)
   
```



