

```{r,echo=FALSE,warning=FALSE, message=FALSE}
source("R/1_global.R")
source("R/2_functions.R")
effort_data<- TRUE
source("R/3_load-and-clean.R")
source("R/4_figures.R")
source("R/5_tables.R")
#source("R/6_analysis-effort.R")
``` 

common<-UBcommon<-c("GN14","GN41","MF","OT16","TLC1","TLC2","TN")
LBcommon<-c("GN18","GN81","TLC4","TLS1","TLS2")


dat$gearC<-as.factor(unlist(lapply(dat$gear, function(x) 
{
  ifelse(any(common==x), x<-paste0(x,"*"),
         ifelse((dat$basin=="LB" && any(LBcommon==x))||(dat$basin=="UB" && x=="TLO1"),x<-paste0(x,"*"),
                unlist(levels(dat$gear))[x])
  )
}
)))
datS<-subset(dat,standard_gear=="yes")




# Quantifying effort and distribution of PSPAP gears
 
## Objectives
In carrying out the goals of the PSPAP, it is important to understand 
the use of gears over space and time. Catch effort by gear is linked 
both to data collection costs (including man-hours for gear deployment), 
as well as power to detect changes in population size. Both costs and 
detection power are important for determining optimal data collection 
strategies for accurate population estimation under budget constraints. 
To study effort by gear type and frequency of gear use, we address the 
following objectives: 


1. Characterize temporal use of gear for PSPAP duration.
2. Quantify effort and associated dispersion for each PSPAP standard 
gear. 


## Methods
For all analyses, the data for the lower and upper basins were 
considered separately. Using the data entries for which effort could be 
calculated (see the following paragraph), the number of deployments for 
each gear was calculated to measure frequency of use. Deployments of 
each gear were considered over time (by year), as well as over space and 
time (by bend within river segment and year). 


The effort, or the time (in minutes) that a gear was set for in a single 
deployment, was calculated using the start and stop times given in the 
PSPAP database. Hence, all data without a start or stop time was 
excluded from this analysis. The relevant data entries from the PSPAP 
database were merged with the gear data. A stop date column was added to 
take into account gears that are set over night, and all data entries 
using times in AM/PM format were converted to 24 hour clock times. 
Effort was calculated as the difference between the stop time and the 
set time in minutes, and negative effort values were removed from the 
analysis. All calculations were done using the statistical computing 
environment R. Any differences in the set date and stop date were taken 
into account by using the `strptime` function in R to link the start and 
stop times to their corresponding set and stop dates, respectively. 


To quantify effort and associated dispersion, probability density 
distributions were fit to the effort data for common gear types by 
basin. We considered a common gear to be a standard gear that was used 
more than a total of 10 times between 2003 and 2016. A gamma 
distribution was fit to the effort data for each common gear in the 
lower basin using the function "fitdistr" from the MASS package in R. 
This process was repeated for each common gear in the upper basin. The 
mean effort, standard deviation of the effort data, minimum effort, 
maximum effort, and median effort were also calculated for each common 
gear by basin. The results from these computations (all done using R) 
are given in Table 2. 


## Results

As expected, results varied by basin. In the lower basin trotlines TLC4 
were deployed the least (24 times) and mini-fike nets were deployed the 
most (7,411 times) out of all common gear types across the PSPAP 
duration (Table 4). In the upper basin, trotlines TLO1 were the least 
deployed common gear type (20 times) and trammel nets were the most 
deployed common gear type (10,924 times) throughout the 14 years 
reported. Gear deployment also varied by year and bend. 



### Gear Use Over Time by Basin
In the lower basin, mini-fyke nets (MF), trammel nets (TN), and 3 types 
of gill nets (GN14, GN18, GN81) were each deployed during every year of 
the PSPAP (see Table 1). Additionally, Table 1 reveals, that a fourth 
gill net (GN41), as well as otter trawls (OT16) and 2 trotline 
configurations (TLC1, TLC2) were used regularly in the lower basin over 
the PSPAP duration. While none of the gears were deployed in the upper 
basin every year of the PSPAP, mini-fyke nets (MF), otter trawls (OT16), 
trammel nets (TN), 2 configurations of gill nets (GN14, GN41), and one 
configuration of trotlines (TLC1) were used consistently. 


Figure 1 nicely illustrates the data in Table 1. Each dot in Figure 1 is 
associated with a gear and year, and represents that at least one 
deployment was made by the indicated gear during the given year. All of 
the deployment dots are connected by lines to illustrate the use of the 
gear over time. Hence, the gears that were used consistently, are those 
that have dots and lines that expand across the majority of the years. 
In the upper basin, Figure 1 shows that the gears used most consistently 
were deployed all but 1 or 2 years of the study with these years most 
likely being near the beginning of the study. 


All of the gears that have been used consistently over the duration of 
the PSPAP have the designation of standard gear. There are, however, 
several configurations of trotlines that were not used very consistently 
over time but are designated as standard gears. In the lower basin there 
are 8 configurations of standard trotlines that were used but not 
consistently (TLC3, TLC4, TLC5, TLC6, TLC7, TLC8, TLS1, TLS2, TLS3) and 
in the upper basin there are 2 configurations of standard trotlines that 
were used inconsistently (TLC2,TL01). 


###Gear Use Over Time by Bend
Gear deployments over time are also important at a smaller spatial 
scale: by bend within river segment. Table 2 gives the number of 
deployments of each gear type within a particular bend during each year 
of the PSPAP. This information is summarized in Table 3. 


No bend was sampled by standard gears during every year of the PSPAP. Two lower basin bends (segment 13 bend 45 and segment 28 bend 5) and three upper basin bends (segment 15 bend 21 and segment 21 bends 2 and 3) were sampled the least at only one deployment of a standard gear in the 14 years of the PSPAP. These lower basin bends were sampled by a trotline (TLS2 & TLC2, respectively) while the upper basin bends were sampled by a trammel net. Overall, the bend with the most total deployments of standard gears over the course of the PSPAP was bend 17 within the upper basin river segment 4 at a total of 719 deployments. In the lower basin, river segment 7 bend 4 had the highest standard gear usage with 368 deployments. 


As seen in Table 3, several standard gears were deployed only once in a given bend over the entire duration of the PSPAP. On the opposite side of the spectrum, in 2005 there were 205 trammel nets deployed within upper basin segment 4 bend 14---the most deployments of a particular standard gear in a river bend in any year. Across years, the most deployments of a single gear type by bend was 503 trammel net deployments, which occurred in upper basin bend 17 within river segment 4. In the lower basin, the bends with the most deployments of a single gear type within and across years were both found in segment 7. Across years, bend 4 had the most gear deployments (127 trotlines, configuration TLC1), while within years bend 17 had the most (32 mini-fyke nets in 2010). 



###Effort Analysis
Effort data (pooled across years and bends) was calculated for each 
common gear by basin (Table 4). In both basins, otter trawls had by far 
the shortest mean effort with an average of 3 minutes per deployment in 
the lower basin and an average of 4 minutes per deployment in the upper 
basin. To compare, 999 minutes (or 16 hours and 39 minutes) per 
deployment by a trotline in the upper basin is the shortest mean effort 
of a common gear when otter trawls are excluded. In both basins, trammel 
nets had the longest mean effort. Trammel nets were used an average of 
1444 minutes (a little over 24 hours) per deployment in the lower basin 
and an average of 1447 minutes per deployment in the upper basin. 


In general, gamma distributions provided a good fit to the effort data. 
The two scenarios that were not fit as well as others, were those where 
the effort data seemed to have multiple modes (LB: TLS1; UB: OT16, TLC1) 
and those where one effort level occurred often and several other effort 
levels occurred rarely, resulting in a probability density function with 
a large spike (LB: TN; UB:GN81, TLC2). 


No bend was sampled by standard gears during every year of the PSPAP.  Two lower basin bends (segment 13 bend 45 and segment 28 bend 5) and three upper basin bends (segment 15 bend 21 and segment 21 bends 2 and 3) were sampled the least at only one deployment of a standard gear in the 14 years of the PSPAP.  These lower basin bends were sampled by a trotline (TLS2 & TLC2, respectively) while the upper basin bends were sampled by a trammel net. Overall, the bend with the most total deployments of standard gears over the course of the PSPAP was bend 17 within the upper basin river segment 4 at a total of 718 deployments.  In the lower  basin, river segment 7 bend 4 had the highest standard gear usage with 368 deployments. 

As seen in Table 3, several standard gears were deployed only once in a given bend over the entire duration of the PSPAP.  On the opposite side of the spectrum, in 2005 there were 204 trammel nets deployed within upper basin segment 4 bend 14---the most deployments of a particular standard gear in a river bend in any year.  Across years, the most deployments of a single gear type by bend was 502 trammel net deployments, which occurred in upper basin bend 17 within river segment 4.  In the lower basin, the most deployments of a single gear type within a year was 32, which occurred in two different bends during different years: bend 17 in segment 7 (32 mini-fyke nets in 2010) and bend 1 in segment 9 (32 TLS1 trotlines in 2007). Across years, segment 7 bend 4 had the most gear deployments (127 trotlines, configuration TLC1).





## Figures
```{r,echo=FALSE}
figures(1)
```

Figure 1. Illustration of gears used since 2003 for Pallid Sturgeon 
Population assessment in the Missouri River. Top panel is for Lower 
basin (RPMA 4) and bottom panel is for Upper basin (RPMA 2). 



## Tables

Table 1. Summary of frequency of use for each gear type by year. An 
asterisk (*) following a gear name indicates that the data from this 
gear was used in the effort analysis (summarized in Table 4). 

```{r,echo=FALSE}
kable(tables(1),
    col.names=c("Basin","PSPAP gear code", "Gear id",2003:2016),
    row.names=FALSE)
   
```


Table 2. Frequency of use (number of deployments) for each standard gear 
type by bend (within river segment) and year. An asterisk (*) following 
a gear name indicates that the data from this gear was used in the 
effort analysis (summarized in Table 4). 

```{r,echo=FALSE}
kable(tables(2),
    col.names=c("Basin","Segment","Bend","PSPAP gear code", "Gear id",2003:2016),
    row.names =FALSE)
   
```

Table 3. Summary of the number of deployments for each standard gear 
type by bend (within river segment) per year. An asterisk (*) following 
a gear name indicates that the data from this gear was used in the 
effort analysis (summarized in Table 4). 

```{r,echo=FALSE}
kable(tables(3),
    col.names=c("Basin","Segment","Bend","PSPAP gear code", "Gear id", "Minimum","Maximum","Mean","Median", "Total Deployments"),
    row.names =FALSE)
   
```

Table 4. Summary of effort data (minutes of use per catch trial) for 
each common gear type. 

```{r,echo=FALSE}
kable(tables(4),
    col.names=c("Basin","PSPAP gear code", "Gear id","No. of Obs.","Mean Effort", "Standard Deviation","Minimum","Maximum", "Median", "Gamma_Shape", "Gamma_Rate"),
    row.names =FALSE)
```



