---
title: "Update"
output: html_document
---

<!--
rmarkdown::render("update.Rmd")
-->


```{r,echo=FALSE,warning=FALSE, message=FALSE}
source("R/1_global.R")
cpue_trnd1<- readRDS("output/cpue_trnd_catchability1e-05_B0sd_fixed_grid.Rds")
summary1<-do.call(rbind, sapply(cpue_trnd1, "[[", "summary", simplify=FALSE))
catch_abund_est<-readRDS("output/catch_abund_catchability_1e-05_B0sd_fixed_grid.Rds")
summary<-do.call(rbind, sapply(catch_abund_est, "[[", "summary", simplify=FALSE))
cpue_trnd4<- readRDS("output/cpue_trnd_catchability4e-05_B0sd_fixed_grid.Rds")
summary4<-do.call(rbind, sapply(cpue_trnd4, "[[", "summary", simplify=FALSE))
``` 


## Figures
```{r,echo=FALSE}
gears<-c("GN14", "GN18", "GN41", "GN81", "MF", "OT16", "TLC1", "TLC2", "TN")
par(mfrow=c(3,3))
for(j in 1:length(gears))
{
  plot(power~B0_sd, data=subset(summary1, gear==gears[j]),main=paste(gears[j]), ylim=c(0,1.2))
}
```

Figure 1. Power of detecting a significant CPUE trend based on 100 replicates for each level of B0_sd with mean catchability set to 0.00001. 

```{r,echo=FALSE}
gears<-c("GN14", "GN18", "GN41", "GN81", "MF", "OT16", "TLC1", "TLC2", "TN")
par(mfrow=c(3,3))
for(j in 1:length(gears))
{
  plot(power~B0_sd, data=subset(summary4, gear==gears[j]),main=paste(gears[j]), ylim=c(0,1.2))
}
```

Figure 2. Power of detecting a significant CPUE trend based on 100 replicates for each level of B0_sd with mean catchability set to 0.00004. 

```{r,echo=FALSE}
gears<-c("GN14", "GN18", "GN41", "GN81", "MF", "OT16", "TLC1", "TLC2", "TN")
par(mfrow=c(3,3))
for(j in 1:length(gears))
{
  plot(bias~B0_sd, data=subset(summary1, gear==gears[j]),main=paste(gears[j]))
}
```

Figure 3. CPUE trend bias from 100 replicates for each level of B0_sd with mean catchability set to 0.00001. 



## Tables

Table 1. Summary of trend data of 100 replicates by gear for B0_sd=0.5 and mean q=0.00001.

```{r,echo=FALSE}
kable(cpue_trnd1[[6]]$summary[c(1:8,11,9,10)],
    col.names=c("Gear","Mean Trend", "Mean SE", "Max SE", "Mean P-value","Max P-value", "Mean Flag Count", "Power", "Bias", "Mean q", "B0 SD"),
    row.names=FALSE)
   
```


Table 2. Summary of abundance estimates of 100 replicates by gear for B0_sd=0.5 and mean q=0.00001.

```{r,echo=FALSE}
kable(catch_abund_est[[6]]$summary,
    col.names=c("Gear","Mean Trend", "Mean SE", "Max SE", "Mean P-value","Max P-value", "Mean Flag Count", "Power", "Bias", "Mean q", "B0 SD"),
    row.names=FALSE)
   
```