---
title: "Estimating bend-level abundance- Mo with axillary information"
author: Michael E. Colvin
date: 4 October 2016
output:
  html_document:
    theme: flatly
    highlight: espresso
    toc: FALSE
---

```{r, echo=FALSE}  
    # RUN SOURCE CODE
    source("./R/1_global.R")
    source("./R/2_functions.R")
    source("./R/3_load.R")
    source("./R/4_clean.R")
    source("./R/5_tables.R")
    source("./R/6_figures.R")   
```

```{r,eval=FALSE}
    source("./R/6_analysis.R")
```
To do:
1. set up rep tracking, sim returns 2 outputs, pit and pit+acoustic for bdn



```{r}
figures(1)
```


<!-- rmarkdown::render("index.Rmd")# -->

# Analysis objectives

1. Evaluate the number of recapture occasions need to estimate capture probability($p$) and abundance ($N$) for
varying population and capture probabilities
2. Evaluate the use of tagged telemetry fish to reduce effort


```{r,echo=FALSE}
library(pander)
source("./R/7_main.R")

```
The design should meet the following criteria:

1. Be estimable at least 90% of simulations 
2. Estimates mean absolute deviation within 20% of true value
3. Mean relative standard error of estimate less than or equal to 20%
    1. Abundance
    2. Demographic rates
        1. Lambda
        2. Survival
        3. Gamma

Segment level estimate
1. closed w/in bend (rd) N = sum(N_bend/bend length)
2. closed w/in segment  
3. 

Rules of thumb, N recaptures? number of fish? 
use shovelnose for capture probabilities?


```{r}

```




