---
title: "Creating a reference population"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: readable
    highlight: tango
---

```{r, echo=FALSE}
source("src/1_global.R")
source("src/2_functions.R")
source("src/3_load-and-clean.R")
source("src/4_figures.R")
source("src/5_tables.R")
```

<!--
rmarkdown::render("index.Rmd",output_format='html_document')# build page

??? 
* How static are habitats within bend... markov matrices of units
* Don't actually measure habitat dimensions? WTF?
-->

# Overview

A reference population of known pallid sturgeon abundance is needed to 
evaluate varying monitoring designs and their ability to achieve the fundamental
objectives identified during the first population assessment program workshop. 


# Objectives

The objectives of this analysis were to:

1. Create a spatially explicit pallid sturgeon population with known
bend-level abundances, and
2. Simulate the reference population for 20 years with varying but known
survival and recruitment


# Methods

## Spatial extent and grain

This analysis was constrained to generate reference pallid sturgeon populations
for segments and bend within Missouri River RPMAs 2 and 4. Bends of varying length
were used to generate the spatial distribution of pallid sturgeon (Table 1).



## Creating a reference population

### Density formulation
We randomly assigned the number of pallid sturgeon occupying each bend given
reported segment level densities by origin (Table 2). The number of pallid sturgeon
assigned to a bend was stochastically generated from a Poisson distribution given the 
segment density accounting for bend length as:

$$N_{segment,bend}\sim Poisson(\lambda_{segment,origin}\cdot Length_{segment,bend}) \text{(1)}$$

where 
 
* $N$ is the number of pallid sturgeon within each bend,
* $\lambda$ is the segment and origin specific density in fish per RKM,
* $Length$ is bend length in RKM,
* $segment$ indexes segment within RPMA,
* $bend$ indexes bend with segment,
* $origin$ indexes whether pallid sturgeon were hatchery or natural origin, and
* $bend$ indexes bend within segment. 

## Abundance formulation

We also used an abundance formulation to provide additional analysis flexibility.
This approach required an segment and origin specific abundance as a vector of 
bend-specific probabilities 


$$N_{segment,bend}\sim Multinomial(Abundance_{segment,origin}, p_{segment,bend}) \text{(2)}$$

where 
 
* $N$ is the number of pallid sturgeon within each bend,
* $Abundance$ is the segment and origin specific total pallid sturgeon abundance,
* $p$ is the bend specific location probability,
* $segment$ indexes segment within RPMA,
* $bend$ indexes bend with segment,
* $origin$ indexes whether pallid sturgeon were hatchery or natural origin, and
* $bend$ indexes bend within segment. 

## Assumptions

1. No movement 
2. No recruitment
3. Survival homogenous among individuals

# Results 

* Number of bends for each RPMA


# Discussion
* Virtual versus power analysis of a post-hoc glm to get trend

How did wildhaber and adam do it did they do virtual or just trend.


# Figures


# Tables


Table 1. Summary of bends within PSPAP Missouri river segments.
 
```{r,echo=FALSE}
tbl<-tables(1)
tbl$basin<- ifelse(tbl$basin=="lower",4,2)
tbl$b_segment<-as.numeric(tbl$b_segment)
tbl<- tbl[order(tbl$b_segment),]
tbl[duplicated(tbl$basin),]$basin<-""
kable(tbl,row.names=FALSE,
    col.names=c("RPMA","Segment","Number of bends",
        "Minimum length (km)", "Mean length (km)","Maximum length (km)"),
    align="lccccc")
```

Table 2. Segment and river level densities by origin


# Code

Box 1. Function used to simulate reference populations given pallid sturgeon
density, survival, and spatial structure. 

```{r,echo=FALSE}
print(reference_population)
```

